---
title: "COVID 19 Tracking"
author: "Juan Ramirez Jr."
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = "")
library(tidyverse)
library(lubridate)
library(geofacet) #install for NY Times Map
library(zoo) #install to deal with time series
library(scales) #can make nicer axis labels for graphs
```

## Setup - United States COVID-19 Cases and Deaths

Here are the NY Times data on cumulative cases by date and state.
```{r}
cases <- read_csv("https://github.com/nytimes/covid-19-data/raw/master/us-states.csv")
```

Note that both `cases` and `deaths` are cumulative by date.

If we want to extract the daily cases we can use the following code which will calculate the difference in (cumulative) cases for one day minus the previous day using the `diff()` function 
```{r}
cases %>%
  group_by(state) %>%
  mutate(case_daily = c(cases[1], diff(cases)),
         death_daily = c(deaths[1], diff(deaths))) %>%
  ungroup()
```


It is also common to report the 7-day moving average to account for reporting delays over the weekends and delays testing results reports.

```{r}
cases %>%
  group_by(state) %>%
  mutate(case_daily = c(cases[1], diff(cases))) %>%
  #filter(!(date == as.Date("2021-03-08") & state == "Missouri")) %>% 
  arrange(date) %>%
  mutate(case_7 = rollmean(case_daily, k = 7, fill = NA)) %>%
  ungroup()
```



Here is data from the US Census Bureau on the state estimated population size. Why might this data be useful?
```{r}
pop <- read_csv("USpop2019.csv")
```

```{r eval = FALSE}
cases %>%
  group_by(state) %>%
  mutate(case_daily = c(cases[1], diff(cases))) %>%
  arrange(date) %>%
  mutate(case_7 = rollmean(case_daily, k = 7, fill = NA, align = "right")) %>%
  ungroup() %>% 
  left_join(pop, by = c("state" = "NAME",
                        "fips" = "GEOID")) %>% 
  write_csv("clean_cases.csv")
```

## Part 1 - Identify Research Questions
Identify 2-3 potential research questions you could try to answer using the provided data (or if you had additional data).

```{r}
cases_avg <- read_csv("clean_cases.csv")
```

Which state had the highest number of positive COVID cases in the year 2020?

Which state had the highest average/daily death rate?

Which state was the safest?

Which state had the most rapid increase in cases?

How do high population states compare to low population states?

What were the worst months of the pandemic?






## Part 2 - Outline the Process
Pick one of the research questions above. Write a brief outline of the steps you would need to take to 1) prepare the data for analysis, and 2) analyze/visualize the data. 

>What were the worst months of the pandemic?

Data Preparation  

- Step 1:  Create (mutate) a new variable to represent the month and year.
- Step 2:  Group by month/year and sum up cases per month or average daily cases per month (either works).


Data Analysis/Visualization  

- Step 1:  
- Step 2:  



## Part 3 - Start Coding
Start to translate your outline into code. You may need to look back at cheatsheets, previous notes, or google items.  Start simple first and then add complication.

```{r}
#data preparation
cases_avg %>% 
  mutate(month_year = lubridate::ymd(date),
         month = month(date, label=TRUE),
         year = year(date)) %>% 
         unite("month_year", month:year, sep = " ", remove = FALSE) %>% group_by(month_year) %>% 
  summarise(total = sum(case_daily, na.rm = TRUE),
            average = mean(case_daily, na.rm = TRUE)) %>% 
  arrange(desc(total)) %>% 
  mutate(month_year_date = my(month_year)) %>% 
  filter(total >= 2576066) %>% 
  ggplot(aes(y = fct_reorder(month_year, total), x = total)) +
  geom_col(aes(fill = factor(year(month_year_date)))) +
  labs(x = "Total Cases Reported", y = "Month and Year",
       title = "Worst Months of the Pandemic",
       subtitle = "Top 10 months by case totals",
       caption = "Source: NY Times",
       fill = "Year") +
  scale_x_continuous(labels = scales::comma) +
  theme_bw()
```

## Part 3.2 - Another example
For this next chunk, I will make a visual to answer the question "Which state had the highest average/daily death rate?"
```{r}
#data preparation
cases_avg %>% 
  group_by(case_daily) %>% 
  filter(case_daily >= 90000) %>% 
  arrange(desc(case_daily)) %>% 
  ggplot(aes(y = case_daily, x = state)) +
  geom_col() +
  labs(x = "State", y = "Daily Cases",
       title = "States with the highest avg daily death rate") +
  theme_bw()
```


```{r}
#data visualization
```


Now take what you have created and think about ways to modify your visualization to make it more interesting or engaging, to clean it up, add details, or highlight details.  

```{r}
#revised visualization
```


## Part 4 - Look Back, Look Ahead
Look back at what you have done, were you able to adequately answer your original research question? Look ahead at what steps you might take next to improve on your analysis or to modify your existing visualization (even if you are not sure how to do it in R).

Yes, I was able to answer both questions quite well!





## Extra Extra
Recently, the NY Times provided a graph of cases rising across the US (scroll down to "After month of progress...")
https://www.nytimes.com/interactive/2021/04/06/us/variants-cases-spread.html


A fellow statistician recreated the graphic using R:
https://livefreeordichotomize.com/2021/04/07/nytimes-map-how-to/

I have already provided the US Census data so you can skip those steps, but give it a try, replicating her code to see if you can recreate the graphic. Comment the code as you determine what each line does.
